Restrictive Clearance Comparison

### 2025-03-20 data-raw/ClearanceProject-2-RelativeLikelihood.Rmd
Ensured in-progress analysis scripts were able to run sequentially. Further refinement expected in the future.

```{r data_loading}
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
all_chem_properties <- readRDS("data-raw/ctxR_3comp2_cheminfo.rds")

# Sipes 2017 data
cyp_data <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"),
  "RestrictiveClearance_Analysis/",
  "CLint_CYP_Sipes2017.csv"
)) %>%
  left_join(
    chem.physical_and_invitro.data %>%
      select(CAS = CAS, Chemical = DTXSID) %>%
      distinct()
  ) %>%
  select(Chemical, CAS, starts_with("MET_")) %>%
  filter(!is.na(Chemical))
names(cyp_data) <- c("Chemical", "CAS", "CYP1A2",
                     "CYP2C9", "CYP2C19", "CYP2D6", "CYP3A4")
# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))
```

@@

```{r venn_diagrams}

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "httk_vs_CvT_VennDiagram_human.png"),
       bg = "white",
       height = 3, width = 5
)

@@

combined_preds <- all_httk_preds %>%
  mutate(Reference = as.character(Reference)) %>%
  left_join(winning_preds) %>%
  filter(!is.na(model))

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 3.8, width = 10
)# Number of chemicals we have RMSLE's for
combined_RMSLE %>% pull(Chemical) %>% unique() %>% length()

@@

# AIC and relative likelihood
# LL from pre-calculated httk predictions
restrictive_ll <- all_httk_preds %>% 
  rename(Conc_est = httk_Preds_restrictive,
         Time_trans = Time) %>%
  select(!httk_Preds_nonrestrictive) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!c(observations, coefs_vector, Time.Units, Time_trans.Units)) %>%
  mutate(model = "model_3comp2_restrictive",
         AIC = -2 * log_likelihood)

nonrestrictive_ll <- all_httk_preds %>% 
  rename(Conc_est = httk_Preds_nonrestrictive,
         Time_trans = Time) %>%
  select(!httk_Preds_restrictive) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!c(observations, coefs_vector, Time.Units, Time_trans.Units)) %>%
  mutate(model = "model_3comp2_nonrestrictive",
         AIC = -2 * log_likelihood) # Simple formula when there are no optimized params

httk_AIC <- bind_rows(restrictive_ll,
                      nonrestrictive_ll,
                      this_AIC %>% filter(model == "model_flat") %>%
                        filter_targets())

@@

this_rel_evaluated_df <- this_css_df %>%
  inner_join(httk_rel_ll %>% distinct(Chemical, Species, ll_label)) %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(this_httk_clearance_df) %>% # CLtot and halflife
  inner_join(data.frame(this_clint_df)) %>%
  inner_join(this_param_df %>% 
               distinct(Chemical,
                        forced_human_values,
                        Fabsgut.restrictive,
                        Fabsgut.nonrestrictive)) %>%
  mutate(Chemical = forcats::fct_reorder(Chemical, restrictive_halflife)) %>%
  ungroup() %>%
  distinct()

halflife_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, restrictive_halflife,
           nonrestrictive_halflife,
           ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = halflife.tkstats, 
                      ymin = nonrestrictive_halflife,
                      ymax = restrictive_halflife),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "halflife (hours)") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

@@

cltot_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, CLtot_Fgutabs.tkstats,
           restrictive_clearance, nonrestrictive_clearance, 
           Fabsgut.restrictive, Fabsgut.nonrestrictive,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = CLtot_Fgutabs.tkstats,
                      ymin = restrictive_clearance/Fabsgut.restrictive,
                      ymax = nonrestrictive_clearance/Fabsgut.nonrestrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "CLtot/Fabsgut") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

@@

```{r functional}
all_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% all_chems, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)
target_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% target_chems, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)

@@

pre_enrichment_mat <- inner_join(
  functional_uses_all %>%
    filter(
      DTXSID %in% all_chems, !(DTXSID %in% target_chems), # all non-targets
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_nontargets"),
  functional_uses_all %>%
    filter(
      DTXSID %in% target_chems, 
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_targets")
) %>%
  janitor::clean_names() %>%
  mutate(n_total_targets = length(target_chems),
         n_total_nontargets = length(all_chems) - n_total_targets,
         n_nontargets_out = n_total_nontargets - n_nontargets,
         n_targets_out = n_total_targets - n_targets) %>%
  select(!c(n_total_nontargets, n_total_targets))

@@

# Makeshift upset plot
upset_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, restrictive_halflife,
           nonrestrictive_halflife,
           ll_label) %>%
left_join(functional_uses_all %>%
            distinct(DTXSID, `Harmonized Functional Use`) %>%
            rename(Chemical = DTXSID) %>%
            filter(`Harmonized Functional Use` %in% c(
              "Biocide",
              "Pharmaceutical (EPA)",
              "Degradant/impurity (EPA)"))
          ) %>%
  # filter(!is.na(`Harmonized Functional Use`)) %>%
  mutate(val = ifelse(is.na(`Harmonized Functional Use`), NA, "Present")) %>%
  distinct(Chemical, ll_label, `Harmonized Functional Use`, val) %>%
  ggplot(aes(x = Chemical, y = `Harmonized Functional Use`, color = val)) + 
  geom_point() +
  scale_color_manual(values = "black") +
  facet_grid(cols = vars(ll_label), scales = "free_x", space = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(color = "white"),
        strip.text = element_blank(),
        legend.title = element_blank(), axis.title.y = element_blank())

@@

# Testing to make sure there are some chemicals that have different clearances
count(this_httk_clearance_df, restrictive_clearance == nonrestrictive_clearance)
# The majority (140) do!

```{r load-other}

# Check to see how much of the full dataset is available
length(intersect(unique(new.raw.httk.data$DTXSID), full_list))

focus_chems <- union(new.raw.httk.data$DTXSID, target_chems)

all_chem_properties %>%
  filter(propertyId %in% c("logkow-octanol-water",
                     "pka-acidic-apparent",
                     "pka-basic-apparent",
                     "water-solubility",
                     "logkoa-octanol-air",
                     "henrys-law",
                     "boiling-point")) %>%
  glimpse()

# Fraction unbound data for target chemicals
fup_df <- this_param_df %>%
  distinct(Chemical, Species, forced_human_values,
           Funbound.plasma, Funbound.plasma.adjustment, 
           Funbound.plasma.dist) %>%
  mutate(
    Funbound.plasma.median = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "^\\d+\\.?\\d+e?-?\\d+(?=,)"
    )),
    Funbound.plasma.lower95 = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "(?<=,)\\d+\\.?\\d+e?-?\\d+(?=,)"
    )),
    Funbound.plasma.upper95 = as.numeric(str_extract(
      Funbound.plasma.dist, pattern = "(?<=,)\\d+\\.?\\d+e?-?\\d$"
    ))
    )

all_hum_rat_fups <- chem.physical_and_invitro.data %>%
  filter(DTXSID %in% all_chems) %>%
  select(DTXSID, Human.Funbound.plasma, Rat.Funbound.plasma) %>%
  mutate(Rat.Funbound.plasma = as.numeric(str_extract(Rat.Funbound.plasma,
                                                      "^([^,]+)")),
         Human.Funbound.plasma = as.numeric(str_extract(Human.Funbound.plasma,
                                                        "^([^,]+)"))
         )

# Get restrictive and non-restrictive parameters

@@

```{r load_Kd}
new.raw.httk.data %>%
  distinct(DTXSID, Affinity.Kd.Med, Affinity.Kd.Low, Affinity.Kd.High,
           Fit) %>%
  na.omit() %>%
  mutate(Target_Chemical = (DTXSID %in% target_chems)) %>%
ggplot(
       aes(
         x = forcats::fct_reorder(DTXSID, Affinity.Kd.Low)
         )) +
  geom_linerange(aes(
    ymin  = Affinity.Kd.Low,
    ymax = Affinity.Kd.High,
    color = Target_Chemical
  ), alpha = 0.4) +
  geom_point(aes(
    y = Affinity.Kd.Med, color = Target_Chemical
  ), size = 0.75) +
  # facet_grid(cols = vars(Fit), space = "free_x", scales = "free_x") +
  scale_y_log10(labels = scales::label_log(digits = 1)) +
  scale_color_manual(values = c("black", "green4")) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "bottom") +
  labs(title = "Kd estimates across chemicals", x = NULL)

@@

kd_v_fup_plot <- new.raw.httk.data %>%
  inner_join(all_hum_rat_fups) %>%
  pivot_longer(cols = Human.Funbound.plasma:Rat.Funbound.plasma,
               names_to = "Species", values_to = "httk_fup_med") %>%
  ggplot(aes(x = Affinity.Kd.Med, y = httk_fup_med, color = Species)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("green4", "purple2"),
                     labels = c("Human", "Rat")) +
  scale_x_log10(labels = scales::label_log(digits = 2)) +
  labs(x = "Dissociation constant Kd\n(uM, from in vitro RED assay)",
       y = "Fraction unbound\nin plasma") +
  theme_bw() +
  theme(legend.position = "left",
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8, color = "black"))

@@

$$
\gamma = \frac{100\cdot \rm{RMSLE}}{\theta}
$$
Where $\theta$ is the relative log-likelihood. The smaller number the better.
If the RMSLE is zero, then it means that it is a "perfect" fit and we are highly
confident in using other metrics to gauge clearance. Below I put these together.

```{r model_weighting}
combined_model_weights <- combined_RMSLE %>%
  left_join(this_rll_probs) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("model_"),
               names_to = "all_models",
               values_to = "relative_likelihood") %>%
  filter(all_models == model) %>%
  ungroup() %>%
  mutate(gamma_weight = 100 * RMSLE_invivoPKfit / relative_likelihood,
         rmsle_comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than")
  ) %>%
  mutate(rmsle_comparison_group = factor(rmsle_comparison_group,
                                         levels = c("Less than", "Between",
                                                    "More than"))) %>%
  select(-all_models) %>% distinct()

@@

```{r combine_tkstats}
# How many overlapping chemicals have winning models that aren't flat?
count(
  this_winmodel,
  model
) # 153 = 97 1-compartment + 56 2-compartment

count(
  this_tkstats,
  # model,
  !is.na(CLtot.tkstats),
  !is.na(CLtot_Fgutabs.tkstats),
  !is.na(halflife.tkstats)
) %>% flextable() %>%
  set_header_labels(
    `!is.na(CLtot.tkstats)` = "Valid Total Clearance",
    `!is.na(CLtot_Fgutabs.tkstats)` = "Valid Total Clearance/Fgutabs",
    `!is.na(halflife.tkstats)` = "Valid halflife",
    n = "Number of Chemicals"
    )
# 83 chemicals have valid total clearance
# 180 chemicals have valid total clearance/Fgutabs
# 5 chemicals have only valid total clearance and no valid total clearance/Fgutabs

count(this_tkstats, is.na(Css.tkstats))

@@

# Table
this_css_df %>%
  count(Css_httk_r == Css_httk_nr,
        Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0,
        Css_httk_r - Css_httk_nr >= 0.1) %>%
  rename("No difference" = "Css_httk_r == Css_httk_nr",
         "Less than 0.1" = "Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0",
         "At least 0.1" = "Css_httk_r - Css_httk_nr >= 0.1")

# Chart with relative differences (think steps) of Css between clearance types
this_css_df %>%
  mutate(relative_difference = (Css_httk_r - Css_httk_nr)/Css_httk_nr) %>%
  count(relative_difference == 0, Css_httk_r == 0) # 31 Chemicals have zero difference
###
this_css_df %>%
  mutate(relative_difference = (Css_httk_r - Css_httk_nr)/Css_httk_nr) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference, .desc = FALSE))) +
  geom_linerange(aes(ymin = Css_httk_nr, ymax = Css_httk_r),
                 linewidth = 1.25) +
  labs(title = "Css differences between restrictive & nonrestrictive clearance",
       y = "Css\n(mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size = 8, color = "black"))

@@

this_css_df %>%
  mutate(relative_difference = log1p((Css_httk_r - Css_httk_nr)/Css_httk_nr)) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference, .desc = FALSE))) +
  geom_linerange(aes(ymin = 0,
                     ymax = 0 + relative_difference),
                 linewidth = 1) +
  labs(
    title = paste0("Relative Css differences between\n",
                   "restrictive & nonrestrictive clearance"),
       y = "log10 Css\n(relative difference, mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size = 8, color = "black"))

@@

```{r clearance_viz}

>>>>

Here I load all the data

```{r data_loading}
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
load("data-raw/rat_pars_rat.RData")
all_chem_properties <- readRDS("data-raw/RCdata/ctxR_3comp2_cheminfo.rds")
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

# Functional use categories from CPdat
functional_uses_all <- readr::read_csv(paste0(
  Sys.getenv("FIG_DIR"), "RestrictiveClearance_Analysis/",
  "ChemExpo_bulk_functional_uses_20250304.csv"
))
```

@@

```{r venn_diagrams}
member_list <- list(cvt = cvt_plasma_chems,
                    httk_3comp2 = human_3compchems)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "httk_vs_CvT_VennDiagram_human.png"),
       bg = "white",
       height = 3, width = 5
)
```

@@

combined_preds <- rph_httk_preds %>%
  left_join(winning_preds) %>%
  filter(!is.na(model))

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 3.8, width = 10
)

# Number of chemicals we have RMSLE's for
combined_RMSLE %>% pull(Chemical) %>% unique() %>% length()

@@

# AIC and relative likelihood
# LL from pre-calculated httk predictions
restrictive_ll <- rph_httk_preds %>% 
  select(!httk_Preds_nonrestrictive) %>%
  rename(Conc_est = httk_Preds_restrictive,
         Time_trans = Time) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!c(observations, coefs_vector, Time.Units, Time_trans.Units)) %>%
  mutate(model = "model_3comp2_restrictive",
         AIC = -2 * log_likelihood)

nonrestrictive_ll <- rph_httk_preds %>% 
  select(!httk_Preds_restrictive) %>%
  rename(Conc_est = httk_Preds_nonrestrictive,
         Time_trans = Time) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!c(observations, coefs_vector, Time.Units, Time_trans.Units)) %>%
  mutate(model = "model_3comp2_nonrestrictive",
         AIC = -2 * log_likelihood) # Simple formula when there are no optimized params

httk_AIC <- bind_rows(restrictive_ll,
                      nonrestrictive_ll,
                      this_AIC %>% filter(model == "model_flat") %>%
                        filter_targets()
                      )

@@

this_rel_evaluated_df <- rph_httk_clearance_df %>%
  inner_join(httk_rel_ll %>% distinct(Chemical, Species, ll_label)) %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(rph_param_df %>% 
               distinct(Chemical,
                        forced_human_values,
                        Fabsgut.restrictive,
                        Fabsgut.nonrestrictive,
                        CLint)) %>%
  mutate(Chemical = forcats::fct_reorder(Chemical, halflife_restrictive)) %>%
  ungroup() %>%
  distinct()

halflife_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, halflife_restrictive,
           halflife_nonrestrictive,
           ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = halflife.tkstats, 
                      ymin = halflife_nonrestrictive,
                      ymax = halflife_restrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "halflife (hours)") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

@@

cltot_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, CLtot_Fgutabs.tkstats,
           CLtot_restrictive, CLtot_nonrestrictive, 
           Fabsgut.restrictive, Fabsgut.nonrestrictive,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = CLtot_Fgutabs.tkstats,
                      ymin = CLtot_restrictive/Fabsgut.restrictive,
                      ymax = CLtot_nonrestrictive/Fabsgut.nonrestrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "CLtot/Fabsgut") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

@@

```{r functional}
all_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% human_3compchems, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)
target_chem_functions <- functional_uses_all %>%
  filter(DTXSID %in% target_chems_human, !is.na(`Harmonized Functional Use`)) %>% 
  group_by(`Harmonized Functional Use`) %>% 
  count() %>% arrange(-n) %>% head(20) %>% pull(1)

@@

pre_enrichment_mat <- inner_join(
  functional_uses_all %>%
    filter(
      DTXSID %in% human_3compchems, !(DTXSID %in% target_chems_human), # all non-targets
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_nontargets"),
  functional_uses_all %>%
    filter(
      DTXSID %in% target_chems_human, 
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_targets")
) %>%
  janitor::clean_names() %>%
  mutate(n_total_targets = length(target_chems_human),
         n_total_nontargets = length(human_3compchems) - n_total_targets,
         n_nontargets_out = n_total_nontargets - n_nontargets,
         n_targets_out = n_total_targets - n_targets) %>%
  select(!c(n_total_nontargets, n_total_targets))

@@

# Makeshift upset plot
upset_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, halflife_restrictive,
           halflife_nonrestrictive,
           ll_label) %>%
left_join(functional_uses_all %>%
            distinct(DTXSID, `Harmonized Functional Use`) %>%
            rename(Chemical = DTXSID) %>%
            filter(`Harmonized Functional Use` %in% c(
              "Biocide",
              "Pharmaceutical (EPA)",
              "Degradant/impurity (EPA)"))
          ) %>%
  # filter(!is.na(`Harmonized Functional Use`)) %>%
  mutate(val = ifelse(is.na(`Harmonized Functional Use`), NA, "Present")) %>%
  distinct(Chemical, ll_label, `Harmonized Functional Use`, val) %>%
  ggplot(aes(x = Chemical, y = `Harmonized Functional Use`, color = val)) + 
  geom_point() +
  scale_color_manual(values = "black") +
  facet_grid(cols = vars(ll_label), scales = "free_x", space = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(color = "white"),
        strip.text = element_blank(),
        legend.title = element_blank(), axis.title.y = element_blank())

@@

# Testing to make sure there are some chemicals that have different clearances
count(rph_httk_clearance_df, CLtot_restrictive == CLtot_nonrestrictive)
# The majority (140) do!

@@

```{r load_Kd}

red_assay_res %>%
  distinct(DTXSID, Affinity.Kd.Med, Affinity.Kd.Low, Affinity.Kd.High,
           Fit) %>%
  na.omit() %>%
  mutate(Target_Chemical = (DTXSID %in% target_chems_human)) %>%
ggplot(
       aes(
         x = forcats::fct_reorder(DTXSID, Affinity.Kd.Low)
         )) +
  geom_linerange(aes(
    ymin  = Affinity.Kd.Low,
    ymax = Affinity.Kd.High,
    color = Target_Chemical
  ), alpha = 0.4) +
  geom_point(aes(
    y = Affinity.Kd.Med, color = Target_Chemical
  ), size = 0.75) +
  # facet_grid(cols = vars(Fit), space = "free_x", scales = "free_x") +
  scale_y_log10(labels = scales::label_log(digits = 1)) +
  scale_color_manual(values = c("black", "green4")) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "bottom") +
  labs(title = "Kd estimates across chemicals", x = NULL)

@@

all_hum_rat_fups <- chem.physical_and_invitro.data %>%
  filter(DTXSID %in% human_3compchems) %>%
  select(DTXSID, Human.Funbound.plasma, Rat.Funbound.plasma) %>%
  mutate(Rat.Funbound.plasma = as.numeric(str_extract(Rat.Funbound.plasma,
                                                      "^([^,]+)")),
         Human.Funbound.plasma = as.numeric(str_extract(Human.Funbound.plasma,
                                                        "^([^,]+)"))
         )

kd_v_fup_plot <- red_assay_res %>%
  inner_join(all_hum_rat_fups) %>%
  pivot_longer(cols = Human.Funbound.plasma:Rat.Funbound.plasma,
               names_to = "Species", values_to = "httk_fup_med") %>%
  ggplot(aes(x = Affinity.Kd.Med, y = httk_fup_med, color = Species)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("green4", "purple2"),
                     labels = c("Human", "Rat")) +
  scale_x_log10(labels = scales::label_log(digits = 2)) +
  labs(x = "Dissociation constant Kd\n(uM, from in vitro RED assay)",
       y = "Fraction unbound\nin plasma") +
  theme_bw() +
  theme(legend.position = "left",
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8, color = "black"))

@@

$$
\gamma = \frac{100\cdot \rm{RMSLE}}{\theta}
$$

Where $\theta$ is the relative log-likelihood. The smaller number the better.
If the RMSLE is zero, then it means that it is a "perfect" fit and we are highly
confident in using other metrics to gauge clearance. Below I put these together.

```{r model_weighting}
combined_model_weights <- combined_RMSLE %>%
  left_join(httk_rel_ll) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("model_"),
               names_to = "all_models",
               values_to = "relative_likelihood") %>% 
  ungroup() %>%
  mutate(gamma_weight = 100 * RMSLE_invivoPKfit / relative_likelihood,
         rmsle_comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than")
  ) %>%
  mutate(rmsle_comparison_group = factor(rmsle_comparison_group,
                                         levels = c("Less than", "Between",
                                                    "More than"))) %>%
  select(-all_models) %>% distinct()

@@

```{r combine_tkstats}
# How many overlapping chemicals have winning models that aren't flat?
count(
  this_winmodel %>% filter_targets(),
  model
) # 153 = 97 1-compartment + 56 2-compartment

count(
  this_tkstats %>% filter_targets(),
  # model,
  !is.na(CLtot.tkstats),
  !is.na(CLtot_Fgutabs.tkstats),
  !is.na(halflife.tkstats)
) %>% flextable() %>%
  set_header_labels(
    `!is.na(CLtot.tkstats)` = "Valid Total Clearance",
    `!is.na(CLtot_Fgutabs.tkstats)` = "Valid Total Clearance/Fgutabs",
    `!is.na(halflife.tkstats)` = "Valid halflife",
    n = "Number of Chemicals"
    )

count(this_tkstats %>% filter_targets(), is.na(Css.tkstats))

@@

# Table
rph_httk_clearance_df %>%
  count(Css_httk_r == Css_httk_nr,
        Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0,
        Css_httk_r - Css_httk_nr >= 0.1) %>%
  rename("No difference" = "Css_httk_r == Css_httk_nr",
         "Less than 0.1" = "Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0",
         "At least 0.1" = "Css_httk_r - Css_httk_nr >= 0.1")

# Chart with relative differences (think steps) of Css between clearance types
rph_httk_clearance_df %>%
  mutate(relative_difference = (Css_httk_r - Css_httk_nr)/Css_httk_nr) %>%
  count(relative_difference == 0, Css_httk_r == 0) # 31 Chemicals have zero difference
###
rph_httk_clearance_df %>%
  mutate(relative_difference = (Css_httk_r - Css_httk_nr)/Css_httk_nr) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference, 
                                      max,
                                      .desc = FALSE))) +
  geom_linerange(aes(ymin = Css_httk_nr, ymax = Css_httk_r),
                 linewidth = 1.25) +
  labs(title = "Css differences between restrictive & nonrestrictive clearance",
       y = "Css\n(mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size = 8, color = "black"))

@@

rph_httk_clearance_df %>%
  mutate(relative_difference = log1p((Css_httk_r - Css_httk_nr)/Css_httk_nr)) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference,
                                      max,
                                      .desc = FALSE))) +
  geom_linerange(aes(ymin = 0,
                     ymax = 0 + relative_difference),
                 linewidth = 1) +
  labs(
    title = paste0("Relative Css differences between\n",
                   "restrictive & nonrestrictive clearance"),
       y = "log10 Css\n(relative difference, mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 10),
        axis.title = element_text(size = 10),
        axis.text.y = element_text(size = 8, color = "black"))

@@

## Exploring _httk_ parameters and solved values for `3compartment2`
